TSELM: Target speaker extraction using discrete tokens and language models

### 2024-09-09 src/paper.tex
update

In Table \ref{hubert_wavlm_6}, we compare the performance of HuBERT and WavLM as SSL model and examine the effects of using either one or multiple layers for discretization. Our findings indicate that using HuBERT as the SSL model results in better DNSMOS scores but much worse dWER compared to our WavLM baseline. The improved DNSMOS scores likely stem from the vocoder performance, yet they do not adequately reflect speech intelligibility, which is crucial for speech separation tasks. 
The poorer dWER scores observed with HuBERT may be attributed to its training on clean speech, which might not equip it to capture the complexity and richness of mixed speech. Moreover, our results from WavLM-L6 suggest that when performing speech separation, discretizing across multiple layers provides better results than relying on a single layer.  This might be because that using multiple layer outputs can better tolerate errors compared to using just one layer output.

@@

\section{Conclusion}
In this work, we introduced a novel way using discrete tokens and language models for target speaker extraction.
Our method leverages multiple hidden layers of WavLM and Kmeans tokenizers for encoding, employs cross-attention and a language model for separation, and utilizes a scalable HiFi-GAN for audio reconstruction.
Experiments have 
shown that our model can achieve excellent performance in terms of speech quality, and comparable performance in terms of speech intelligibility and 
speaker similarity. However, we still have a gap between continuous methods especially in speech intelligibility and speaker similarity. Future research should focus on shrinking 
this gap.

\section{Acknowledgment}
We want to thank Super Computing Network (SCNet) for providing the computing resources.

>>>>

In Table \ref{hubert_wavlm_6}, we compare the performance of HuBERT and WavLM as SSL model and examine the effects of using either one or multiple layers for discretization. Our findings indicate that using HuBERT as the SSL model results in better DNSMOS scores but much worse dWER compared to our WavLM baseline. The improved DNSMOS scores likely stem from the vocoder performance, yet they do not adequately reflect speech intelligibility, which is crucial for speech separation tasks. SkIM-UniCATS uses only DNSMOS as evaluation, which might not be enough 
in terms of speech intelligibility and speaker similarity.
The poorer dWER scores observed with HuBERT may be attributed to its training on clean speech, which might not equip it to capture the complexity and richness of mixed speech. Moreover, our results from WavLM-L6 suggest that when performing speech separation, discretizing across multiple layers provides better results than relying on a single layer.  This might be because that using multiple layer outputs can better tolerate errors compared to using just one layer output.

@@

\section{Conclusion}
In this work, we introduce a novel way using discrete tokens and language models for target speaker extraction.
Our method leverages multiple hidden layers of WavLM and Kmeans tokenizers for encoding, employs cross-attention and a language model for separation, and utilizes a scalable HiFi-GAN for audio reconstruction.
Experiments have 
shown that our model can achieve excellent performance in terms of speech quality, and comparable performance in terms of speech intelligibility and 
speaker similarity. However, we still have a gap between continuous methods especially in speech intelligibility and speaker similarity. Future research should focus on shrinking 
this gap.

\section{Acknowledgment}
We woudld like to thank Super Computing Network (SCNet) for providing the computing resources.
